\documentclass[11pt]{report}

\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{enumitem}
\usepackage{url}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}

\title{Personal Notes on Jordan Quillen (Co)homology}
\author{}
\date{\today}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[chapter]
\newtheorem{example}{Example}[chapter]
\newtheorem{note}{Note}[chapter]
\newtheorem{question}{Question}[chapter]

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[chapter]
\newtheorem{lemma}{Lemma}[chapter]

\begin{document}
\maketitle
\tableofcontents

\input{chapters/01-background}
\input{chapters/02-workspace-architecture}
\input{chapters/03-experiments}
\input{chapters/04-experiment-summary}
\input{chapters/05-theory-roadmap}

\chapter{Current Repository Assessment}

\section{Snapshot Boundary}

This chapter records a working assessment of the repository state. It is a
private planning note, not a claim file and not proof evidence. Any statement
below that depends on generated results should be checked against the relevant
\path{results.json}, \path{experiment.md}, commit record, and verification
commands before it is quoted in public project documentation.

\begin{note}
The current project stage is best described as
\[
  \text{reproducible experimental scaffold}
  \quad+\quad
  \text{small exact computations}.
\]
It is not yet a completed theory of intrinsic Jordan Quillen homology. The
useful achievement is narrower but important: the repository now has enough
notation, claims, tests, and toy computations to expose false shortcuts and to
shape the next definitions.
\end{note}

\section{Overall Diagnosis}

The project goal is clear. The intended construction is intrinsic: use the
Jordan algebra category itself, rather than first replacing a Jordan algebra by
an associated Lie algebra such as the Tits--Kantor--Koecher algebra. The guiding
objects are derived abelianization, Jordan cotangent complexes, low-degree
Quillen cohomology, square-zero extensions, and comparisons with Glassman or
TKK-style constructions.

The repository discipline is also good. It asks for a strict distinction
between proved claims, proof drafts, conjectures, and computations. It also
requires exact arithmetic where possible and forbids fabricated experiment
output. This is especially important here because small Jordan examples can
look convincing even when a naive relation complex is not homotopically
invariant.

At the code level, the package is intentionally light: a Python package over
Python \(3.11\) or later, with no runtime dependencies and development tools
such as \path{pytest}, \path{ruff}, and \path{pre-commit}. The \path{Makefile}
exposes entries such as \path{test}, \path{lint}, \path{claims}, and
\path{experiments}. A current practical caveat is that \path{make experiments}
does not mean that every experiment is rerun: the top-level
\path{scripts/run_all_experiments.py} currently behaves as a conservative
status/reporting surface and prints that experiments are not run by default.

\section{Claim Status}

The theory layer remains deliberately conservative.

\begin{itemize}[leftmargin=*]
  \item \path{CLAIM-0001} is draft-level: zero-multiplication Jordan algebras
  are the expected abelian objects in the relevant nonunital setting.
  \item \path{CLAIM-0002} is draft-level: the indecomposables functor
  \(Q(J)=J/J^2\) is expected to be the abelianization or left adjoint in the
  chosen absolute nonunital setting.
  \item \path{CLAIM-0003} is conjectural: Beck modules over Jordan algebras
  should match a precise Jordan module convention, but the identity extraction
  still needs to be completed.
  \item \path{CLAIM-0004} is conjectural: universal Jordan differentials
  \(\Omega_{B/A}\) need a settled coefficient category and universal property.
  \item \path{CLAIM-0005} is conjectural: the transitivity triangle for the
  Jordan cotangent complex is not yet proved in the project.
\end{itemize}

\section{Experiment Status Matrix}

The current experiment landscape is:

\begin{description}[leftmargin=*,style=nextline]
  \item[001 square-zero, run.]
  A finite-grid sanity check for one-dimensional split extensions.
  \item[002 two-dimensional, not run.]
  Representative small examples still need to be computed.
  \item[003 Beck modules, not run.]
  Symbolic module identities are still a placeholder target.
  \item[004 TKK comparison, not run.]
  Comparison examples have not yet been computed.
  \item[005 derived \(Q\), run.]
  The one-generator toy benchmark matches the expected dimensions.
  \item[006 \(J_3\) stress test, run.]
  Naive Presentation B has an extra false \(H_1\) class.
  \item[007 two-step repair, run.]
  One hand-coded degree \(2\) generator kills that false class in the tested
  low-degree complex.
\end{description}

The status should be read from each experiment's own files, especially
\path{results.json} or \path{results.md}. The global runner is not yet a
complete experiment execution engine.

\section{Experiment 001: Split Square-Zero Extensions}

Experiment 001 studies one-dimensional split square-zero extensions
\[
  E = J\ltimes M,
  \qquad
  M^2=0.
\]
It tests two one-dimensional bases:
\[
  e^2=e
  \qquad\text{and}\qquad
  e^2=0,
\]
with \(M=km\) and scalar action
\[
  L_e(m)=\lambda m.
\]

The important methodological point is that the experiment uses exact rational
arithmetic and checks the Jordan identity on a finite grid of linear
combinations, not only on basis vectors. This catches a real failure mode:
\(\lambda=2\) can pass a basis-only smoke test but fail once general linear
combinations are tested.

For \(e^2=e\), the tested nonunital split extension passes for
\[
  \lambda\in\{0,\tfrac12,1\}.
\]
If \((e,0)\) is required to remain a unit in the extension, the surviving value
is \(\lambda=1\). For \(e^2=0\), the tested action is forced to
\[
  \lambda=0.
\]

This experiment supports the square-zero workflow behind \path{CLAIM-0003},
but it does not prove the Beck-module classification. Its main lesson is that
future module-identity checks must use symbolic derivation or a strong
polynomial identity check, not basis-vector testing alone.

\section{Experiment 005: One-Generator Derived Indecomposables}

Experiment 005 is the cleanest current Quillen-style toy computation. It
studies one-generator presentations
\[
  F(x),\quad
  F(x)/(x^2-x),\quad
  F(x)/(x^2),\quad
  F(x)/(x^3),\quad
  F(x)/(x^4).
\]

The toy rule is that after applying indecomposables \(Q\), only the linear part
of a relation enters the first relation complex. In other words, the current
toy complex is the two-term complex
\[
  k^{\{\text{relations}\}}
  \longrightarrow
  k^{\{\text{generators}\}},
\]
where the differential records relation linear terms. Exact row-rank
computations then give \(H_0\) and \(H_1\).

The current benchmark conclusions are:
\[
\begin{array}{c|c}
\text{presentation} & (H_0,H_1) \\
\hline
F(x) & (1,0) \\
F(x)/(x^2-x) & (0,0) \\
F(x)/(x^2) & (1,1) \\
F(x)/(x^3) & (1,1) \\
F(x)/(x^4) & (1,1).
\end{array}
\]

This gives a useful low-degree intuition for \(Q(J)=J/J^2\): a relation with a
linear term can kill \(H_0\), while a pure higher-order relation appears as a
naive \(H_1\) class in this one-step model. The boundary remains essential:
this is not a general cofibrant replacement engine and it does not establish
presentation invariance.

\section{\texorpdfstring{Experiment 006: Presentation Stress Test For \(J_3\)}{Experiment 006: Presentation Stress Test For J3}}

Experiment 006 compares two presentations of
\[
  J_3=(t)/(t^3).
\]
The first presentation is
\[
  \text{A: } F(x)/(x^3),
\]
and the second is
\[
  \text{B: } F(x,y)/(y-x^2,\;xy,\;y^2).
\]

Presentation A agrees with the one-generator toy computation and gives
\[
  H_0=1,\qquad H_1=1.
\]
The naive one-step diagnostic for Presentation B gives
\[
  H_0=1,\qquad H_1=2.
\]
The result is therefore not a failure of the project. It is a useful negative
test: the naive relation complex sees an extra false \(H_1\) class for
Presentation B. Without higher syzygy data or a genuine cofibrant replacement,
Presentation B cannot be used as evidence for presentation invariance.

\begin{note}
The practical target created by Experiment 006 is precise: identify the extra
class in the naive \(H_1\), then attach a higher cell or syzygy generator that
kills it without changing the intended algebra.
\end{note}

\section{\texorpdfstring{Experiment 007: Two-Step Toy Repair For \(J_3\)}{Experiment 007: Two-Step Toy Repair For J3}}

Experiment 007 performs exactly that fixed repair. It starts from Presentation
B and adds one degree \(2\) generator \(b\), designed to kill the redundant
class represented after applying \(Q\) by \(a_3\), the relation corresponding
to \(y^2=0\).

The low-degree indecomposable complex has the form
\[
  Q_2=\langle b\rangle
  \longrightarrow
  Q_1=\langle a_1,a_2,a_3\rangle
  \longrightarrow
  Q_0=\langle x,y\rangle,
\]
with
\[
  d_Q(a_1)=y,\qquad
  d_Q(a_2)=0,\qquad
  d_Q(a_3)=0,\qquad
  d_Q(b)=a_3.
\]
Hence \(b\) kills exactly the extra class that was visible in the naive
Presentation B computation.

The computed dimensions become
\[
  H_0=1,\qquad H_1=1,\qquad H_2=0.
\]
Thus the naive \(H_1\)-dimension \(2\) from Experiment 006 drops to \(1\), in
agreement with Presentation A through the tested degree.

This is the most important current mathematical evidence in the repository:
the project has moved from detecting a presentation artifact to hand-coding a
low-degree semi-free repair that removes it. The result must still be quoted
carefully. It is a fixed \(J_3\) computation, not an automatic syzygy search,
not a general cofibrant replacement algorithm, and not a proof of full
presentation invariance.

\section{Current Strengths}

The main strengths of the current repository are:

\begin{itemize}[leftmargin=*]
  \item The research boundary is explicit: computation is evidence, not proof.
  \item The claim files are conservative and do not overstate the theory.
  \item The package already has exact linear algebra utilities for small
  examples.
  \item The indecomposables code computes the rank of \(J^2\) and therefore
  \(\dim Q(J)=\dim J-\dim J^2\) in finite examples.
  \item The tests cover basic cases such as zero multiplication, a
  one-dimensional idempotent algebra, and a one-dimensional square-zero
  algebra.
  \item Experiments 006 and 007 give a concrete warning and repair pattern for
  presentation-sensitive toy computations.
\end{itemize}

\section{Current Risks}

The main risks are:

\begin{enumerate}[leftmargin=*]
  \item The coefficient theory is not yet settled. The files for modules,
  universal differentials, and TKK comparison remain draft or placeholder
  surfaces rather than completed theory.
  \item Experiment 007 is hand-coded. It should inspire the next algorithm, but
  it is not itself an automatic cell-attachment or syzygy-discovery procedure.
  \item Experiments 002, 003, and 004 are still the missing bridge between the
  one-generator toy computations and the broader Jordan algebra theory.
  \item The global experiment runner does not yet summarize the whole
  experiment state in a machine-checkable matrix.
\end{enumerate}

\section{Next Experimental Priorities}

\subsection*{Priority 1: Experiment 008}

The next best experiment is an automatic syzygy-discovery version of
Experiment 007. The goal is to input the three first-stage relations
\[
  a_1,\quad a_2,\quad a_3
\]
from Presentation B and automatically find the degree \(2\) attachment whose
image under \(Q\) kills \(a_3\). The expected output should reproduce
\[
  H_0=1,\qquad H_1=1,\qquad H_2=0,
\]
while explicitly saying that this is only a degree \(1/2\) toy replacement, not
a full cofibrant replacement.

\subsection*{Priority 2: Experiment 003}

The Beck-module identity experiment should be restored and made symbolic. The
starting point should be the Jordan identity
\[
  ((a^2b)a)=a^2(ba)
\]
inside a square-zero extension \(J\ltimes M\) with \(M^2=0\). The experiment
should extract the \(M\)-linear terms and compare the resulting operator
identities for \(L_x\) with the classical Jordan module axioms. This is central
for \path{CLAIM-0003}.

\subsection*{Priority 3: Experiment 002}

The two-dimensional example table should be completed in a modest,
representative way before attempting classification. Good first examples are:
the two-dimensional zero product algebra, \(ke\oplus km\) with \(e^2=e\), a
one-dimensional idempotent plus a zero ideal, and a small two-dimensional
nilpotent example. For each example, record \(\dim J^2\), \(\dim Q(J)\),
derivation dimensions, and candidate relations for \(\Omega\).

\subsection*{Priority 4: Experiment 004}

The TKK comparison should begin with two or three minimal examples and should
allow disagreement as a valid outcome. The repository rule is correct: one
should not identify intrinsic Jordan Quillen homology with TKK Lie homology
without either a proof or a counterexample.

\subsection*{Priority 5: Experiment Status Aggregator}

The script \path{scripts/run_all_experiments.py} should become a true
read-only status aggregator before it becomes an execution driver. A useful
first version would read each \path{experiment.md} and \path{results.json} or
\path{results.md}, then produce a matrix recording status, pass/fail data, and
the presentation-invariance flags that already appear in the JSON outputs. By
default it should avoid rewriting results.

\section{Working Conclusion}

The most important current mathematical progression is
\[
  \text{Experiment 006}
  \quad\longrightarrow\quad
  \text{Experiment 007}.
\]
Experiment 006 shows that a naive Presentation B relation complex for \(J_3\)
contains an extra false \(H_1\) class. Experiment 007 shows that a specific
two-step semi-free toy replacement kills that class and restores the expected
low-degree dimensions through degree \(1\).

The next phase should therefore not jump directly to paper-level theorems. It
should first turn the hand-coded success of Experiment 007 into a repeatable
low-degree method, while also completing the coefficient-theory and small
example experiments that make the theory meaningful.

\appendix
\input{chapters/appendix-notation-and-literature}

\end{document}
