\chapter{Experiment Summary}

\section{What The Current Runs Support}

The current executed experiments support the usefulness of exact arithmetic
small examples. Experiment 001 checks square-zero extension behavior in tiny
families, Experiment 005 validates a one-generator toy rule, and Experiment 006
shows why presentation issues cannot be ignored. Experiment 007 then repairs
the fixed \(J_3\) Presentation B diagnostic through degree \(1\) by adding one
explicit degree \(2\) generator.

\section{What The Current Runs Do Not Prove}

The successful toy runs do not prove the full Jordan Quillen homology theory.
They do not establish a general cofibrant replacement construction, a
presentation-invariance theorem, or an equivalence between intrinsic Jordan
Quillen homology and TKK Lie homology.

\section{Most Important Boundary}

The most important current warning is the pair of Experiments 006 and 007.
Experiment 006 shows that a naive one-step relation complex for Presentation B
has an extra false \(H_1\) class. Experiment 007 shows that this particular
false class can be killed by a hand-coded degree \(2\) generator, but the
repair is still example-specific and low-degree.

\section{Personal Takeaway}

The experiments are best used as guardrails for definitions. They are useful
when they expose impossible claims, missing hypotheses, or misleading
computational shortcuts. The 007 run is especially useful as a template for
what later automatic low-degree cell attachment would need to reproduce.
