\chapter{Theory Roadmap}

\section{Abelianization}

The first theoretical target is to make \(Q(J)=J/J^2\) precise in the chosen
category. The related claim files are currently draft-level and should not be
quoted as checked theorems.

\section{Beck Modules And Square-Zero Extensions}

The next target is to decide whether Beck modules over a Jordan algebra match
the classical Jordan module identities. This is central because it controls the
coefficient category for cohomology.

\section{Derivations And Universal Differentials}

After the coefficient category is clear, the project should construct and test
universal Jordan differentials \(\Omega_{B/A}\). The universal property remains
a conjectural or proof-draft target until the module convention is settled.

\section{Cotangent Complex And Low Degrees}

The cotangent complex \(L^{\mathrm{Jord}}_{B/A}\) requires a resolution
framework, such as simplicial Jordan algebras, dg Jordan algebras over the
Jordan operad, or a comparison between both.

Experiment 007 suggests the next concrete computational target: replace naive
one-step relation complexes by a principled low-degree cell-attachment
procedure. The present evidence is only the fixed \(J_3\) two-step repair, so
it should guide implementation tests rather than be quoted as a theorem.

Low-degree cohomology should eventually explain \(D^0\) in terms of Jordan
derivations and \(D^1\) in terms of square-zero extensions. These statements
need separate claim files before they become paper-level assertions.

\section{Comparison Problems}

The comparison with Glassman cohomology and TKK Lie-theoretic constructions
should remain explicit and cautious. Agreement in small examples is useful,
but non-agreement may be equally informative.
